---
title: "Chapter 4 - The Ames Housing Data"
format: 
    html:
        embed-resources: true
---


```{r setup}
use("tidyverse")
use("tidymodels")
tidymodels_prefer()

data(ames)
```



```{r plotting}

glimpse(ames)

# median(ames$Sale_Price)
summary(ames$Sale_Price)

first <- 
    ames |>
    ggplot(aes(x = Sale_Price)) +
    geom_histogram(bins = 50, col = "white")

first
```

excerpt from book:
    - This plot shows us that the data are right-skewed; there are more inexpensive houses than expensive ones.
    -  When modeling this outcome, a strong argument can be made that the price should be log-transformed
        - pros
            - advantages of this type of transformation are that no houses would be predicted with negative sale prices 
            - errors in predicting expensive houses will not have an undue influence on the model
            -  from a statistical perspective, a logarithmic transform may also stabilize the variance in a way that makes inference more legitimate
        - cons 
            - The disadvantages of transforming the outcome mostly relate to interpretation of model results.

```{r log_transform}
first + scale_x_log10()

ames <- 
    ames |> 
    mutate(Sale_Price = log10(Sale_Price))
```


```{r ames-to-parq}
#| eval: false

nanoparquet::write_parquet(
    ames,
    file = fs::path_wd("data", "ames", ext = "parquet"),
    compression = "gzip"
)

```